---
title: "Data  Challenge 1 - Yucheng Liu"
output: html_document
date: "2023-09-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r }
## load in the required libraries 
library(readr)
library(tidyverse)
library(here)
```

```{r}
## read in the data using here() function
state_milk_production <- read_csv(here("data/state_milk_production.csv"))
clean_cheese <- read_csv(here("data/clean_cheese.csv"))
fluid_milk_sales <- read_csv(here("data/fluid_milk_sales.csv"))
milk_products_facts <- read_csv(here("data/milk_products_facts.csv"))
milkcow_facts <- read_csv(here("data/milkcow_facts.csv"))

## disable scientific notation
options(scipen = 999)
```

## Figure 1.
Make a scatter plot showing the milk production in pounds versus the average number of milk cows using milkcow_facts. Write a caption for this plot.

```{r}
# Create the scatter plot
p1 <- ggplot(milkcow_facts, aes(x = avg_milk_cow_number, y = milk_production_lbs, color = year)) +
  geom_point() +
  labs(
    title = "Milk Production vs Number of Milk Cows",
    x = "Number of Milk Cows (in millions)",
    y = "Milk Production (in millions of pounds)",
    color = "Year"
  ) +
  theme_minimal()
p1 <- p1 + labs(
  caption = "Figure 1: This plot shows the relationship between the number of dairy cows (in millions) and milk production \n
  (in millions of pounds) from the 1980s to the 2010s. Before 2000, the number of dairy cows was large but the \n
  yield was always low, the milk yield has been on the rise. Although the number of dairy cows increased very \n
  little after 2000, milk production increased significantly and peaked around 2010. The plot shows a rough \n
  linear relationship, indicating that as the number of milk cows decreases, milk production increases over time."
)
# Print the plot
print(p1)
```

## Figure 2.
Examine the milk total milk production in each region over time using state_milk_production. We will plot line plots for the milk production versus the year faceted by region. Write a caption for this plot.

```{r}
# Summarise the total milk production by year and region
state_milk_production_summarised <- state_milk_production %>%
  group_by(year, region) %>%
  summarise(total_milk_production = sum(milk_produced))

# Create the line plots
p2 <- ggplot(state_milk_production_summarised, aes(x = year, y = total_milk_production, color = region)) +
  geom_line() +
  facet_wrap(~region) +
  labs(
    title = "Total Milk Production by Region Over Time",
    x = "Year",
    y = "Total Milk Production (in millions of pounds)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

p2 <- p2 + labs(
  caption = "Figure 2: This line graph shows the total milk production by region over time. The lines for the Appalachian,\n 
  Corn Belt, Delta States, Lake States, Northern Plains, and Pacific regions show an overall increase in milk production \n 
  over time. The lines for the Mountain, Northeast, and Southeastregions show a decrease in milk production over time."
)

# Print the plot
print(p2)
```

## Figure 3
Make a line plot showing the amount (in pounds) of each milk products sold over time using fluid_milk_sales. Write a caption for this plot.

```{r}
# Create the line plot
p3 <- ggplot(fluid_milk_sales, aes(x = year, y = pounds, color = milk_type)) +
  geom_line() +
  scale_y_log10() +
  labs(
    title = "Milk Product Sales Over Time",
    x = "Time",
    y = "Pounds of Product (Log Scale)",
    color = "Milk Product"
  ) +
  theme_minimal()
# Add a caption to the plot
p3 <- p3 + labs(
  caption = "Figure 3: This line plot shows the sales of different milk products over time. The lines \n
  represent different milk products: buttermilk, eggnog, flavored (not whole), low fat (1%), \n
  reduced fat (2%), skim, total production, and whole. The lines show that sales of \n
  buttermilk, eggnog, and flavored (not whole) milk products have decreased over time, while \n
  sales of low fat (1%), reduced fat (2%), skim, and whole milk products have increased."
)

# Print the plot
print(p3)
```

## Figure 4
Make a stacked bar plot of the different type of cheeses consumed over time using a long format version of clean_cheese that you create. Write a caption for this plot.

```{r}
clean_cheese_long <- clean_cheese %>%
  select(-contains("Total")) %>%
  pivot_longer(cols = -Year, names_to = "Cheese_Type", values_to = "Pounds4") %>%
  mutate(Cheese_Type = tolower(Cheese_Type),
         Cheese_Type = str_to_title(Cheese_Type))

# Create the stacked bar plot
p4 <- ggplot(clean_cheese_long, aes(x = Year, y = Pounds4, fill = Cheese_Type)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Set3") +
  labs(x = "Time", y = "Pounds of Cheese") +
  ggtitle("Cheese Consumption Over Time") +
  theme_minimal() 

p4 <- p4 + labs(
  caption = "“Figure 4: The stacked bar plot 'Cheese Consumption Over Time' illustrates the consumption trends \n
  of various types of cheese over time. Each color represents a different type of cheese, and \n 
  the height of each color segment within a bar indicates the quantity consumed in that time  \n
  period. The plot provides insights into both overall cheese consumption and preferences for  \n 
  different types of cheese.”"
)

print(p4)
```

```{r}
milk_products_facts_long <- milk_products_facts %>%
  select(-contains("fluid_milk")) %>%
  select(-contains("cheese")) %>%
  select(-contains("frozen")) %>%
  pivot_longer(cols = -year, names_to = "Milk_Type", values_to = "Pounds") %>%
  mutate(Milk_Type = str_to_title(Milk_Type))

my_palette <- c("#000000", "#FF0000", "#00FF00", "#0000FF", "#FFFF00", "#00FFFF", "#FF00FF",
                "#C0C0C0", "#808080", "#800000", "#808000", "#008000", "#800080", "#008080",
                "#000080", "#666699", "#003366", "#6699CC", "#CC3366", "#99CC33")

p5 <- ggplot(milk_products_facts_long, aes(x = year, y = Pounds, fill = Milk_Type)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer("my_palette") +
  labs(
    title = "Milk Consumption Over Time",
    x = "Time",
    y = "Pounds of Milk"
  ) +
  theme_minimal()

p5 <- p5 + labs(
  caption = "Figure 4: This stacked bar plot, illustrates the consumption trends of various types \n
  of milk products over time. Each color represents a different type of milk product, \n
  and the height of each color segment within a bar indicates the quantity consumed in \n 
  that time period. The x-axis represents time (in years), and the y-axis represents the \n 
  pounds of milk consumed. The plot provides insights into both overall milk \n 
  consumption and preferences for different types of milk products."
)

print(p5)
```



