---
title: "Lab #1: Using dplyr and ggplot2 to recreate a plot"
date: "02/10/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries  

```{r, message = FALSE}
install.packages("ggpubr")
library(ggpubr)
library(tidyverse)
library(medicaldata)
```

## Plot A  

Make a stacked bar plot of the test results by day  
Change the colors using the following hex code colors: #1f77b4, #e377c2 and #2ca02c  
Change the labels on the x-axis, y-axis, title, and legend  
Use theme_minimal() to remove the gray background  
Use theme() to center the title  
Use geom_segment() and annotate() to draw the arrow and write text on the plot  

```{r}
covid <- covid_testing %>%
  select(pan_day, result) %>%
  mutate(result = str_to_title(result)) # Making custom data set to plot in A

# This is part A of the submission

plot_a <- covid %>%
  ggplot(aes(x = pan_day,
             fill = result)) +
  geom_bar() +    # Bar plot
  theme_minimal() +
  scale_fill_manual(values = c("Positive" = "#2ca02c",
                               "Negative" = "#e377c2",
                               "Invalid" = "#1f77b4")) +    # Custom colors
  labs(x = "Pandemic Day",     # Labeling
       y = "Total Number of Tests",
       title = "COVID-19 Test Results at CHOP: \n First 107 Days of the Pandemic") +
  theme(plot.title = element_text(hjust = 0.5,
                                  size = 15),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10)) +
  guides(fill = guide_legend(title = "Test Result")) +    # Legend renaming
  geom_segment(aes(x = 68,   # Arrow
                   y = 230,
                   xend = 78.8,
                   yend = 20),
               linewidth = 1,
               arrow = arrow(type = "open", length = unit(0.5, "cm"))) +
  annotate("text", x = 65, y = 230 + 40,
           label = "Highest Number of \n Positive Tests (20) \n on Day 79",
           size = 3)        # Annotation
```

## Plot B  

Use the covid_summary dataframe  
Make a bar plot and change the colors of your bar plot  
Using scale_y_continuous() change the y-axis labels to percents  
Change the labels on the x-axis, y-axis, and title  
Use theme_minimal() to remove the gray background  
Use theme() to center the title and remove the legend

```{r}
# This is part B of the submission

# Filter the data for COVID tests
covid_summary <- covid_testing %>%
  filter(test_id == "covid") %>%
  # Convert the result to title case for consistency
  mutate(result = stringr::str_to_title(result))

# Group the data by pandemic day and calculate the positive rate
covid_summary <- covid_summary %>%
  group_by(pan_day) %>%
  summarise(positive_rate = sum(result == "Positive") / n()) # Calculate the positive rate

# Create a bar plot of positive rate by pandemic day
plot_b <- ggplot(covid_summary, aes(x = pan_day, y = positive_rate)) +
  # Use identity stat for bars to represent actual values
  geom_bar(stat = "identity", fill = "#2ca02c") +
  # Scale y-axis to percent format
  scale_y_continuous(labels = scales::percent) +
  # Add labels for axes and title
  labs(x = "Pandemic Day", y = "Percent Positive tests", title = "Percent of Positive \n Test Results by Day", lwd = 2) +
  # Use a minimal theme for a clean look
  theme_minimal() +
  # Center the title and adjust its size
  theme(plot.title = element_text(hjust = 0.5, size = 11), 
        # Remove the legend as it's not necessary for this plot
        legend.position = "none",
        # Adjust the size of axis titles
        axis.title.x = element_text(size = 11),           
        axis.title.y = element_text(size = 11))

```

## Plot C  
Use the covid_ages dataframe  
Make a bar plot and change the colors of your bar plot  
Use theme_minimal() to remove the gray background  
Use theme() to center the title, remove the legend, and rotate the x-axis labels

```{r}
# This is part C of the submission

## Read in the covid testing data
covid <- medicaldata::covid_testing
## Filter the data for COVID tests and start with capital letters
covid_test <- filter(covid, test_id == "covid") %>%
  mutate(result = stringr::str_to_title(result))
## Calculates the percent of positive tests per day 
covid_summary <- group_by(covid_test, pan_day) %>%
summarise(total = sum(result == "Positive"))

covid_ages <- covid_test %>%
## Filter for only positive test results
  filter(result == "Positive") %>%
## create a variable for age categories
  mutate(age_categories = cut(age, breaks = c(-Inf, 10, 20, 30, 40, 50, 60, Inf), 
                              labels = c("0-10", "11-20", "21-30", "31-40", "41-50", "51-60", "61+")))

## Initialize ggplot object
plot_c <- ggplot(covid_ages, 
       aes(age_categories)) +
## Bar plot in green
  geom_bar(fill = "#2ca02c") +
## Relabel x and y axis
  labs(x = "Age", y = "Number of Positive Tests", 
## Add a title
       title = "Positive Test by Age Group") +
## Remove the gray background
  theme_minimal() +
## Adjust text size and title position
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.x = element_text(size = 15),           
        axis.title.y = element_text(size = 15),            
        legend.position = "none")
```

## Putting the plots together  
Use ggpubr::ggarrange to arrange the plots  
Label the plots “A”, “B”, and “C”  
Create a common legend and put this at the bottom of the plot

```{r}

# Putting everything together and final output

## First row with plot a
ggarrange(plot_a,               
## Second row with plot b and c
  ggarrange(plot_b, plot_c, ncol = 2, 
## Relabel plot b and c
            labels = c("B", "C")),
            nrow = 2, 
## Relabel plot a
  labels = "A" ,
## Add a common legend at the bottom
  common.legend = TRUE, legend = "bottom")
```
